\bigskip GRASS GIS Vector State of the Art -- Gearing towards GRASS GIS 7 \bigskip

### 2014-04-21 grassposter/2014_EGU_G7_Landscape/poster.tex
add some text to 'data sources' box
git-svn-id: https://svn.osgeo.org/grass/grass-promo@59866 15284696-431f-4ddb-bdfa-cd5b030d7da7
updated poster
git-svn-id: https://svn.osgeo.org/grass/grass-promo@59867 15284696-431f-4ddb-bdfa-cd5b030d7da7

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Abstract}{
% \small <<- too small for a poster!
The upcoming GRASS GIS 7 release improves not only raster processing and general design but the vector processing in the first place. GRASS GIS, as a topological GIS, recognizes that the topology plays the key role in the vector processing and analysis.\newline
Topology ensures that adjacent geographic components in a single vector map are related. In contrast to non-topological GIS, a border common to two areas exists only once and is shared between the two areas. Topological representation of vector data helps to produce and maintain vector maps with clean geometry as well as enables the user to perform certain analyses that can not be conducted with non-topological or spaghetti data. Non-topological vector data are automatically converted to a topological representation upon import. Further more, various cleaning tools exist to remove non-trivial topological errors.\newline
In the upcoming GRASS GIS 7 release the vector library was particularly improved to make it faster and more efficient with an improved internal vector file format. This new topological format reduces memory and disk space requirements, leading to a generally faster processing. Opening an existing vector requires less memory providing additionally support for large files. The new spatial index performs queries faster (compared to GRASS GIS 6 more than 10 times for large vectors). As a new option the user can select a file-based version of the spatial index for large vector data. All topological cleaning tools have been optimized with regard to processing speed, robustness, and system requirements.\newline
The vector engine comes with a new prototype for direct read/write support of OGR Simple Features API.\newline
Additionally vector data can be directly exchanged with topological PostGIS 2 databases. This enables GRASS to read and write topological primitives beside native file-based format also to the topological PostGIS 2 databases. \newline
Considering the wide spread usage of Esri Shapefile, a non-topological format for vector data exchange, it is particularly advantageous that GRASS GIS 7 offers advanced cleaning tools.\newline
For power users and programmers, the new Python interface allows to directly access functions provided by the underlying C libraries; this combines the ease of writing Python scripts with the power of optimized C functionality in the library backend.
}

@@

The following \textbf{basic topological elements} can be edited directly: point, centroid, line, and boundary. A GRASS vector map can contain a~combination of several different types of the elements.
From these basic geometry types the following \textbf{derived topological elements} can be generated: area (closed ring of boundaries + centroid), isle (closed ring of boundaries, no centroid), and node (at both ends of lines/boundaries). Isles and Nodes are not visible to the user. Furthermore face, kernel (3D centroid) and volume (3D area) as defined in the format.

% TODO: point to http://grass.osgeo.org/programming7/

% TODO: 2. explain topology changes <<- needed?

% TODO: show the internal format change or rather the actual model?
\begin{center}
 \includegraphics[width=0.48\textwidth]{odp_slides/grass7_basic_vector_types_digitizer}
 \hspace{10mm}
 \includegraphics[width=0.48\textwidth]{odp_slides/grass7_derived_vector_types}
 \newline
 Figure 1a: Basic and derived topological elements in GRASS GIS 7
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Data Sources}{
This
\begin{center}
 \includegraphics[width=0.47\textwidth]{images/grass-data-sources}
  \newline
 Figure 2: Data sources
\end{center}
}

@@

@@

\blocknode{Lidar}{
\smallskip
The Lidar library ({\url {www.liblas.org}}) used by GRASS GIS permits the import of LAS Lidar
data. The imported data can be in raster ({\bf r.in.lidar} using statistics of choice) or in vector format ({\bf v.in.lidar}). \newline
On-farm water storage study with lidar data in NSW (Australia) developed a full remote sensing monitoring methodology
of water availability with lidar-based bathymetric survey and multi-source remote sensing survey [8].\newline
\begin{center}
 %\includegraphics[width=0.4\textwidth]{images/ofs1}
 %\newline
 %Figure 5: On-Farm-Water-Storage Lidar survey and Depth-Volume-Area surveying [8]

 \includegraphics[width=0.4\textwidth]{images/grass7_las_support}
 \newline
 Figure 4: Example for LAS support in GRASS GIS 7: rapid LAS data assessment through binning
\end{center}
}

\blocknode{Other Improvements \& Additions}{
\smallskip

\begin{itemize}
 \item {\bf v.distance:} Calculates distances from points, lines, or areas to points, lines, or areas.
 \item {\bf v.overlay:} The processing speed has been substantially improved.
 \item {\bf v.net.*:} All vector network analysis tools provide now fine control over node costs.
 \item {\bf v.voronoi:} New option to create Voronoi diagrams for areas.
 \item {\bf v.rectify:} Vector data can now be georeferenced using various methods for 2D and 3D coordinate transformation.
\end{itemize}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{References}{
% \scriptsize <<- too small for a poster!
\begin{center}
\begin{tabular}{rp{0.8\textwidth}}
[Neteler 2012] & Neteler \& Bowman \&  Landa \& Metz, 2012. Environment \& Modeling Software, 31:124-130\\{}
[Gebbert 2014] & Gebbert \& Pebesma, 2014. TGRASS: A temporal GIS for field based environmental modeling, Environmental Modelling \& Software 53:1–12.\\{}
[Zambelli 2013] & Zambelli \& Gebbert \& Ciolli, 2013. PyGRASS: An Object Oriented Python API for GRASS GIS. ISPRS International Journal of Geo-Information 2.1:201-219.\\{}
[Neteler 2005] & Neteler \& Grasso \& Michelazzi \& Miori \& Merler \& Furlanello, 2005. International Journal of Geoinformatics, 1(1):51-61.\\{}
[Landa 2013] & Landa, 2013. Vektorová architektura systému GRASS GIS [GRASS GIS Vector Architecture]. PhD thesis, CTU in Prague, Czech Republic.
\end{tabular}
\end{center}
\smallskip
\hrulefill
\vspace{-5pt}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Abstract}{
% \small <<- too small for a poster!
The upcoming GRASS GIS 7 release improves not only raster processing and general design but the vector processing in the first place. GRASS GIS, as a topological GIS, recognizes that the topology plays the key role in the vector processing and analysis.\newline\linebreak
Topology ensures that adjacent geographic components in a single vector map are related. In contrast to non-topological GIS, a border common to two areas exists only once and is shared between the two areas. Topological representation of vector data helps to produce and maintain vector maps with clean geometry as well as enables the user to perform certain analyses that can not be conducted with non-topological or spaghetti data. Non-topological vector data are automatically converted to a topological representation upon import. Further more, various cleaning tools exist to remove non-trivial topological errors.\newline\linebreak
In the upcoming GRASS GIS 7 release the vector library was particularly improved to make it faster and more efficient with an improved internal vector file format. This new topological format reduces memory and disk space requirements, leading to a generally faster processing. Opening an existing vector requires less memory providing additionally support for large files. The new spatial index performs queries faster (compared to GRASS GIS 6 more than 10 times for large vectors). As a new option the user can select a file-based version of the spatial index for large vector data. All topological cleaning tools have been optimized with regard to processing speed, robustness, and system requirements.\newline\linebreak
The vector engine comes with a new prototype for direct read/write support of OGR Simple Features API.\newline
Additionally vector data can be directly exchanged with topological PostGIS 2 databases. This enables GRASS to read and write topological primitives beside native file-based format also to the topological PostGIS 2 databases. \newline
Considering the wide spread usage of Esri Shapefile, a non-topological format for vector data exchange, it is particularly advantageous that GRASS GIS 7 offers advanced cleaning tools.\newline\linebreak
For power users and programmers, the new Python interface allows to directly access functions provided by the underlying C libraries; this combines the ease of writing Python scripts with the power of optimized C functionality in the library backend.
}

@@

% TODO: point to http://grass.osgeo.org/programming7/
% TODO: 2. explain topology changes <<- needed?
% TODO: show the internal format change or rather the actual model?
\begin{center}
 \includegraphics[width=0.48\textwidth]{odp_slides/grass7_basic_vector_types_digitizer}
 \hspace{10mm}
 \includegraphics[width=0.48\textwidth]{odp_slides/grass7_derived_vector_types}
 \newline
 Figure 1a: Basic and derived topological elements in GRASS GIS 7
\end{center}

The following \textbf{basic topological elements} can be edited directly: point, centroid, line, and boundary. A GRASS vector map can contain a~combination of several different types of the elements.
From these basic geometry types the following \textbf{derived topological elements} can be generated: area (closed ring of boundaries + centroid), isle (closed ring of boundaries, no centroid), and node (at both ends of lines/boundaries). Isles and Nodes are not visible to the user. Furthermore face, kernel (3D centroid) and volume (3D area) as defined in the format.\newline

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Data Sources}{

  GRASS GIS 7 supports various data sources in read and write
  access. Beside native file-based raster and vector format GRASS
  allows to specify external data sources through GDAL/OGR
  library. GRASS GIS 7 also natively supports PostGIS database
  including its topological extension.

\begin{center}
 \includegraphics[width=0.47\textwidth]{images/grass-data-sources}
  \newline
 Figure 2: Supported data sources
\end{center}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{Other Improvements \& Additions}{
\smallskip

\begin{itemize}
 \item {\bf v.distance:} Calculates distances from points, lines, or areas to points, lines, or areas.
 \item {\bf v.overlay:} The processing speed has been substantially improved.
 \item {\bf v.net.*:} All vector network analysis tools provide now fine control over node costs.
 \item {\bf v.voronoi:} New option to create Voronoi diagrams for areas.
 \item {\bf v.rectify:} Vector data can now be georeferenced using various methods for 2D and 3D coordinate transformation.
\end{itemize}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\blocknode{Lidar}{
\smallskip
The Lidar library ({\url {www.liblas.org}}) used by GRASS GIS permits the import of LAS Lidar
data. The imported data can be in raster ({\bf r.in.lidar} using statistics of choice) or in vector format ({\bf v.in.lidar}). \newline
\begin{center}
 \includegraphics[width=0.7\textwidth]{images/grass7_las_support}
 \newline
 Figure 4: Example for LAS support in GRASS GIS 7: rapid LAS data assessment through binning
\end{center}

On-farm water storage study with lidar data in NSW (Australia) developed a full remote sensing monitoring methodology of water availability with lidar-based bathymetric survey and multi-source remote sensing survey [Chemin 2011].\newline
\begin{center}
\includegraphics[width=0.4\textwidth]{images/ofs1}
\hspace{5mm}
\includegraphics[width=0.3\textwidth]{images/ofs2}
\newline
Figure 5: On-Farm-Water-Storage Lidar survey and Depth-Volume-Area surveying [8]
\end{center}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\blocknode{References}{
% \scriptsize <<- too small for a poster!
\begin{center}
\begin{tabular}{rp{0.8\textwidth}}
[Chemin 2011] & Chemin \& Rabbani, 2011. International Journal of Geoinformatics,  7(3):1-6.\\{}
[Gebbert 2014] & Gebbert \& Pebesma, 2014. TGRASS: A temporal GIS for field based environmental modeling, Environmental Modelling \& Software 53:1–12.\\{}
[Landa 2013] & Landa, 2013. Vektorová architektura systému GRASS GIS [GRASS GIS Vector Architecture]. PhD thesis, CTU in Prague, Czech Republic.\\{}
[Neteler 2005] & Neteler \& Grasso \& Michelazzi \& Miori \& Merler \& Furlanello, 2005. International Journal of Geoinformatics, 1(1):51-61.\\{}
[Neteler 2012] & Neteler \& Bowman \&  Landa \& Metz, 2012. Environment \& Modeling Software, 31:124-130\\{}
[Zambelli 2013] & Zambelli \& Gebbert \& Ciolli, 2013. PyGRASS: An Object Oriented Python API for GRASS GIS. ISPRS International Journal of Geo-Information 2.1:201-219.\\{}
\end{tabular}
\end{center}
\smallskip
\hrulefill
\vspace{-5pt}
